### 2023-01-18 analysis/decisionEntropy.Rmd
Measuring a tendency to make decisions according to strict rules

---
title: "Decision entropy"
author: "Matan Mazor"
date: "2023-01-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE)
library(groundhog)

groundhog.library(c(
  'png',
  'grid',
  'ggplot2',
  'svglite',
  'xtable',
  'papaja',
  'tidyverse',
  'broom',
  'cowplot',
  'MESS', # for AUCs
  'lsr', # for effect sizes
  'pwr', # for power calculations
  'brms', # for mixed effects modeling
  'BayesFactor', # for Bayesian t test
  'jsonlite', # parsing data from sort_trial
  'caret' #for cross validation
), '2022-12-01')

load('../.Rdata')
```

## Decision entropy

Unlike the main analysis, where we look at the likelihood of hitting a ship, here we look at the prior of clicking on a cell, regardless of possible ships' locations.

```{r get-priors, echo=FALSE, cache=TRUE}

E2.prior_from_group <- E2.click_log_with_boards %>% 
  filter(test_part=='nonpretend')%>%
  mutate(flat_position = factor(flat_position, levels=seq(25))) %>%
  group_by(grid_number,click_number, flat_position) %>%
  summarise(n=n())%>%
  group_by(click_number, flat_position)%>%
  do( data.frame(.,other_n=sapply(.$grid_number, function(i) mean(.$n[!.$grid_number %in% i])))) %>%
  group_by(grid_number, click_number) %>%
  mutate(p=other_n/sum(other_n)) %>%
  dplyr::select(grid_number,click_number,flat_position,p) %>%
  spread(flat_position,p, fill=0, sep='_') %>%
  ungroup();

```

```{r decision-entropy-functions, echo=FALSE, cache=TRUE}

get_group_prior <- function(grid_i, click_i, board_state) {

  board_state = scan(text= board_state, what = numeric(), sep="," , quiet = TRUE);

  hidden_cells = ifelse(is.na(board_state),1,0);

  group_prior = E2.prior_from_group %>%
    filter(grid_number==grid_i & click_number==click_i) %>%
    select(flat_position_1:flat_position_25) %>%
    as.numeric()

  group_prior_for_hidden_cells = group_prior*hidden_cells/sum(group_prior*hidden_cells)

  return(paste(group_prior_for_hidden_cells,collapse=','))

};

get_prior_click <- function(prior, flat_position) {
  prior <- scan(text= prior, what = numeric(), sep="," , quiet = TRUE);
  return(prior[flat_position])
}

get_prior_click_rank <- function(prior, flat_position) {
  prior <- scan(text= prior, what = numeric(), sep="," , quiet = TRUE);
  rank_prior = rank(-prior)
  return(rank_prior[flat_position])
}

get_prior_entropy <- function(prior) {
  prior <- scan(text= prior, what = numeric(), sep="," , quiet = TRUE);
  prior[which(prior==0)]=NA; #to avoid infinity*0
  entropy <- -sum(unlist(prior)*(unlist(log(prior))),na.rm=TRUE);
  return(entropy);
}

```

```{r p_click, echo = FALSE, cache=TRUE}

# E2.click_log_with_boards <-E2.click_log %>%
#   dplyr::select(subj_id,genuine_first,test_part,grid_number,click_number,i,j,RT,hit_bin) %>%
#   mutate(flat_position = 1+i*5+j) %>%
#   group_by(subj_id,genuine_first,test_part,grid_number) %>%
#   arrange(click_number) %>%
#   summarise(click_number = click_number,
#             board_state=create_board_states(flat_position,hit_bin),
#             flat_position=flat_position,
#             i=i,
#             j=j,
#             RT=RT,
#             hit_bin=hit_bin) %>%
#   rowwise()%>%
#   mutate(likelihood = get_likelihood(board_state),
#          posterior = get_posterior(likelihood),
#          p_click = get_p_click(posterior,flat_position),
#          entropy = get_posterior_entropy(posterior),
#          p_click_rank = get_p_click_rank(posterior,flat_position));

E2.click_log_with_boards <- readRDS('../../experiments/Battleships2/data/batch1/click_log_with_boards.Rda')

get_random_flat_position <- function(board_state) {

  board_state = scan(text= board_state, what = numeric(), sep="," , quiet = TRUE);
  return(which(is.na(board_state))%>%sample(1))

}

get_optimal_flat_position <- function(posterior) {

  posterior = scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
  return(which.max(posterior))

}

E2.click_log_with_boards_optimal <- E2.click_log_with_boards %>%
  mutate(flat_position = get_optimal_flat_position(posterior),
         p_click = get_p_click(posterior,flat_position),
         p_click_rank = get_p_click_rank(posterior,flat_position),
         test_part = 'optimal')

E2.click_log_with_boards_random <- E2.click_log_with_boards %>%
  mutate(flat_position = get_random_flat_position(board_state),
         p_click = get_p_click(posterior,flat_position),
         p_click_rank = get_p_click_rank(posterior,flat_position),
         test_part = 'random')

E2.mean_P_click_rank <- E2.click_log_with_boards %>%
  rbind(E2.click_log_with_boards_optimal) %>%
  rbind(E2.click_log_with_boards_random) %>%
  group_by(subj_id,test_part, genuine_first) %>%
  summarise(p_click_rank=mean(p_click_rank));

E2.mean_P_click_rank_misses_only <- E2.click_log_with_boards %>%
  rbind(E2.click_log_with_boards_optimal) %>%
  rbind(E2.click_log_with_boards_random) %>%
  rowwise() %>%
  filter(!hit_bin) %>%
  group_by(subj_id,test_part, genuine_first) %>%
  summarise(p_click_rank=mean(p_click_rank))

p<- E2.mean_P_click_rank %>%
    filter(test_part != 'optimal') %>%
    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
    ggplot(aes(x=p_click_rank,fill=test_part))+
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
    labs(x='p(ship in clicked square) - rank', y='number of players')+
    theme(legend.position=c(0.10,0.85)) +
    # facet_wrap(~genuine_first, nrow=2)+
    # scale_x_reverse() +
    theme_classic();

ggsave('../figures/E2_pclick_rank.pdf',p)
ggsave('../figures/E2_pclick_rank.png',p,width=5,height=2,dpi=300)

p<- E2.mean_P_click_rank_misses_only %>%
    filter(test_part != 'optimal') %>%
    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
    ggplot(aes(x=p_click_rank,fill=test_part))+
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
    labs(x='p(ship in clicked square) - rank', y='number of players')+
    scale_x_continuous(limits=c(0,15))+
    theme(legend.position=c(0.10,0.85)) +
    # facet_wrap(~genuine_first, nrow=2)+
    # scale_x_reverse() +
    theme_classic();

ggsave('../figures/E2_pclick_rank_misses_only.pdf',p)
ggsave('../figures/E2_pclick_rank_misses_only.png',p,width=5,height=2,dpi=300)

E2.num_irrational_clicks <- E2.click_log_with_boards %>%
  group_by(subj_id,test_part, genuine_first) %>%
  summarise(nic=sum(p_click==0)/5);

p<- E2.num_irrational_clicks %>%
    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
    ggplot(aes(x=nic,fill=test_part))+
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    labs(x='mean number of irrational clicks per game', y='number of players')+
    theme(legend.position=c(0.10,0.85)) +
    facet_wrap(~genuine_first, nrow=2)+
    theme_classic();

ggsave('figures/../E2_num_irrational_clicks.png',p,width=5,height=3,dpi=300)

 E2.RT_p_click_cor <- E2.click_log_with_boards %>%
     filter(click_number<9 & click_number>1) %>%
     group_by(subj_id,test_part, genuine_first) %>%
     summarise(r=cor(RT,p_click,method='spearman'));

 E2.RT_entropy_cor <- E2.click_log_with_boards %>%
     filter(click_number<9 & click_number>1) %>%
     group_by(subj_id,test_part, genuine_first) %>%
     summarise(r=cor(entropy,p_click,method='spearman'))

```
## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.