### 2023-01-27 analysis/shipCompletion.Rmd

```{r ship-completion}

get_hit_number <- function(hit) {

  num_A <- 0;
  num_B <- 0;
  num_C <- 0;

  hit_number <- c();

  for (i in seq(length(hit))) {
    if (hit[i]==0) {
      hit_number = c(hit_number,NaN)
    } else if (hit[i]=='A') {
      num_A = num_A + 1;
      hit_number = c(hit_number,num_A)
    } else if (hit[i]=='B') {
      num_B = num_B + 1;
      hit_number = c(hit_number,num_B)
    } else if (hit[i]=='C') {
      num_C = num_C + 1;
      hit_number = c(hit_number,num_C)
    }
  }

  return(hit_number)
}

revealed_three <- function(hit_number) {

  revealed_three <- c();
  three = FALSE;

  for (i in seq(length(hit_number))) {
    revealed_three = c(revealed_three, three);
     if (!(is.na(hit_number[i])) & hit_number[i]==3) {
      three = TRUE
    } 
  }

  return(revealed_three)
}

E2.hit_order <- E2.click_log %>%
  group_by(genuine_first,subj_id,test_part,grid_number) %>%
  summarise(click_number = click_number,
            hit = hit,
            RT = RT,
            hit_number = get_hit_number(hit),
            revealed_three = revealed_three(hit_number),
            hit_type = ifelse(hit_number==1, 'first',
                              ifelse(hit_number==2 & !revealed_three, 'second',
                                     ifelse(hit_number==2 & revealed_three, 'secondlast',
                                            ifelse(hit_number==3,'thirdlast','sea'))))) %>%
  mutate(prev_hit_type = lag(hit_type),
         prev_hit = lag(hit))

E2.RT_by_hit_order <- E2.hit_order %>%
  group_by(genuine_first,subj_id,test_part,prev_hit_type) %>%
  summarise(RT=median(RT))

p <- E2.RT_by_hit_order %>%
  filter(!is.na(prev_hit_type))%>%
  mutate(prev_hit_type = factor(prev_hit_type,levels=c('first','second','secondlast','thirdlast'), labels = c('first','second','second and last','third')))%>%
  group_by(test_part,prev_hit_type) %>%
  summarise(RT = median(RT),
            sem_RT = bootstrap_error(RT,N_perm)) %>%
  ggplot(aes(x=prev_hit_type,y=RT,fill=test_part)) +
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  scale_color_manual(values=c("#69b3a2", "#404080")) +
  geom_bar(stat='identity',color='black',position=position_dodge()) +
  geom_errorbar(aes(ymin=RT-sem_RT,ymax=RT+sem_RT),width=.2,position=position_dodge(.9))+
  labs(fill=' ',x='previous hit', y='median RT')+
  theme_classic()

ggsave('../docs/figures/E2_RT_by_previous_hit_number_collapse.png',p,width=5,height=2,dpi=300)

##correct for trial order

E2.corrected_RT_by_hit_order <- E2.hit_order %>%
  group_by(subj_id,test_part,click_number) %>%
  mutate(corrected_RT = RT-mean(RT))%>%
  rowwise() %>%
  group_by(genuine_first,subj_id,test_part,prev_hit_type) %>%
  summarise(corrected_RT=median(corrected_RT))

p <- E2.corrected_RT_by_hit_order %>%
  filter(!is.na(prev_hit_type))%>%
  mutate(prev_hit_type = factor(prev_hit_type,levels=c('first','second','secondlast','thirdlast'), labels = c('first','second','second and last','third')))%>%
  group_by(test_part,prev_hit_type) %>%
  summarise(sem_corrected_RT = se(corrected_RT,N_perm),
            corrected_RT = median(corrected_RT)) %>%
  ggplot(aes(x=prev_hit_type,y=corrected_RT,fill=test_part)) +
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  scale_color_manual(values=c("#69b3a2", "#404080")) +
  geom_bar(stat='identity',color='black',position=position_dodge()) +
  geom_errorbar(aes(ymin=corrected_RT-sem_corrected_RT,ymax=corrected_RT+sem_corrected_RT),width=.2,position=position_dodge(.9))+
  labs(fill=' ',x='previous hit', y='median RT, corrected for serial order')+
  theme_classic()

ggsave('../docs/figures/E2_corrected_RT_by_previous_hit_number_collapsed.png',p,width=5,height=2,dpi=300)

## correct for true ship size

E2.RT_by_hit_order_boats_only <- E2.hit_order %>%
  filter(prev_hit!='A')%>%
  group_by(genuine_first,subj_id,test_part,prev_hit_type) %>%
  summarise(RT=median(RT))

p <- E2.RT_by_hit_order_boats_only %>%
  filter(!is.na(prev_hit_type))%>%
  mutate(prev_hit_type = factor(prev_hit_type,levels=c('first','second','secondlast','thirdlast'), labels = c('first','second','second and last','third')))%>%
  group_by(test_part,prev_hit_type) %>%
  summarise(RT = median(RT),
            sem_RT = bootstrap_error(RT,N_perm)) %>%
  ggplot(aes(x=prev_hit_type,y=RT,fill=test_part)) +
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  scale_color_manual(values=c("#69b3a2", "#404080")) +
  geom_bar(stat='identity',color='black',position=position_dodge()) +
  geom_errorbar(aes(ymin=RT-sem_RT,ymax=RT+sem_RT),width=.2,position=position_dodge(.9))+
  labs(fill=' ',x='previous hit', y='median RT')+
  theme_classic()

ggsave('../docs/figures/E2_corrected_RT_by_previous_hit_number_collapsed.png',p,width=5,height=2,dpi=300)

E2.corrected_RT_by_hit_order_boats_only <- E2.hit_order %>%
  filter(prev_hit!='A')%>%
  group_by(subj_id,test_part,click_number) %>%
  mutate(corrected_RT = RT-mean(RT))%>%
  rowwise() %>%
  group_by(genuine_first,subj_id,test_part,prev_hit_type) %>%
  summarise(corrected_RT=median(corrected_RT))

p <- E2.corrected_RT_by_hit_order_boats_only %>%
  filter(!is.na(prev_hit_type))%>%
  mutate(prev_hit_type = factor(prev_hit_type,levels=c('first','second','secondlast','thirdlast'), labels = c('first','second','second and last','third')))%>%
  group_by(test_part,prev_hit_type) %>%
  summarise(sem_corrected_RT = se(corrected_RT,N_perm),
            corrected_RT = median(corrected_RT)) %>%
  ggplot(aes(x=prev_hit_type,y=corrected_RT,fill=test_part)) +
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  scale_color_manual(values=c("#69b3a2", "#404080")) +
  geom_bar(stat='identity',color='black',position=position_dodge()) +
  geom_errorbar(aes(ymin=corrected_RT-sem_corrected_RT,ymax=corrected_RT+sem_corrected_RT),width=.2,position=position_dodge(.9))+
  labs(fill=' ',x='previous hit', y='median RT, corrected for serial order')+
  theme_classic()

ggsave('../docs/figures/E2_corrected_RT_by_previous_hit_number_collapsed_boats_only.png',p,width=5,height=2,dpi=300)

```