### 2023-01-27 docs/half_games.rmd

```{r}

plot_posterior <- function(posterior,file_name) {
  posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
  board_df <- data.frame(i=rep(c(1,2,3,4,5),each=5),j=rep(c(1,2,3,4,5),5),posterior=posterior);
  p <- ggplot(board_df,aes(x=j,y=i,fill=posterior)) +
    geom_tile() +
    coord_equal() +
    scale_x_continuous(breaks=c()) +
    scale_y_continuous(breaks=c(),trans='reverse') +
    scale_fill_gradient(low='#649bc8',high='#c83237', na.value ='white')+
    labs(x='',y='') +
    theme_classic();

  ggsave(paste('figures/',file_name,'.png',sep=''),width=5,dpi=300);
    ggsave(paste('figures/',file_name,'.svg',sep=''),width=5,dpi=300);

}

boardA <- "NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,1,1,0,0,NA,NA,0,NA,NA,1,1,NA,NA,NA"
posteriorA <- boardA%>%get_likelihood()%>%get_posterior();
plot_posterior(posteriorA,'posteriorA');

boardB <- "NA,NA,NA,NA,0,NA,NA,0,NA,NA,NA,NA,NA,NA,0,NA,NA,1,0,NA,NA,NA,1,NA,1"
posteriorB <- boardB%>%get_likelihood()%>%get_posterior();
plot_posterior(posteriorB,'posteriorB');

```

```{r}

E2.half_game_click_df <- E2.df %>%
  dplyr::select(subj_id, 
         test_part, 
         grid_number, 
         num_clicks,
         click_log,
         genuine_first) %>%
  mutate(click_log = gsub("\'","\"", click_log)) %>%
  filter(test_part=='pretend_half_game' | test_part=='nonpretend_half_game');

E2.half_game_click_log <- data.frame(matrix(ncol=10,nrow=0, 
                               dimnames=list(NULL, 
                                             c("subj_id",
                                              "test_part", 
                                              "grid_number", 
                                              "num_clicks",
                                              "i",
                                              "j",
                                              "hit",
                                              "t",
                                              "click_number",
                                              "genuine_first"))))

for (row in 1:nrow(E2.half_game_click_df)) {

    subject_click_log <- data.frame(fromJSON(E2.half_game_click_df[row, ]$click_log)) %>%
    mutate(
      click_number = 1:n(),
      subj_id = E2.half_game_click_df[row, ]$subj_id,
      test_part = E2.half_game_click_df[row, ]$test_part,
      grid_number = E2.half_game_click_df[row, ]$grid_number,
      num_clicks = E2.half_game_click_df[row, ]$num_clicks,
      genuine_first = E2.half_game_click_df[row, ]$genuine_first
    )

    E2.half_game_click_log <- rbind(E2.half_game_click_log, subject_click_log);
}

E2.half_game_click_log <- E2.half_game_click_log %>%
  relocate(subj_id, .before = i) %>%
  relocate(genuine_first, .before = i) %>% 
  relocate(test_part, .before=i) %>% 
  relocate(grid_number, .before=i) %>% 
  relocate(click_number, .before=i) %>%
  group_by(subj_id,test_part,grid_number) %>%
  mutate(RT=t-lag(t,default=0)) %>%
  filter(!(subj_id %in% E2.lucky_pretenders))

```

```{r}

E2.half_game_first_clicks <- E2.half_game_click_log %>%
  filter(click_number==1) %>%
  group_by(i,j, grid_number,test_part) %>%
  summarise(n=n()) %>%
  ungroup() %>%
  complete(i,j,grid_number,test_part,fill=list(n=0)) %>%
  group_by(grid_number,test_part) %>%
  mutate(n=n/sum(n), x=i*5+j);

p<- E2.half_game_first_clicks %>%
  ggplot(aes(x=j,y=i,fill=n)) +
  geom_tile(color='black') +
  # geom_text(aes(label=printnum(n)))+
  coord_equal() +
  scale_x_continuous(breaks=c()) +
  scale_y_continuous(breaks=c(),trans='reverse') +
  scale_fill_gradient(low='#649bc8',high='#c83237', na.value ='white')+
  labs(x='',y='') +
  theme_classic() + 
  facet_wrap(~grid_number+test_part, nrow=2);

ggsave('figures/E2_half_game_first_clicks.png',width=10,height=10,dpi=300)
ggsave('figures/E2_half_game_first_clicks.svg',width=10,height=10,dpi=300)

p<- E2.half_game_first_clicks %>%
  group_by(grid_number,test_part) %>%
  mutate(p=n/sum(n)) %>%
  ggplot(aes(x=j,y=i)) +
  geom_tile(color='black',fill = NA) +
  # geom_text(aes(label=printnum(n)))+
  coord_equal() +
  # Area fills, convert factors to numeric
  geom_rect(
    aes(xmin = as.numeric(j) - 0.5, 
        xmax = as.numeric(j) + 0.5,
        ymin = as.numeric(i) + 0.5-p,
        ymax = as.numeric(i) + 0.5),
    fill='black',alpha=1

  )+
  scale_x_continuous(breaks=c()) +
  scale_y_continuous(breaks=c(),trans='reverse') +
  labs(x='',y='') +
  theme_classic() + 
  facet_wrap(~grid_number+test_part, nrow=2);

ggsave('figures/E2_half_game_first_clicks_height.svg',width=10,height=10,dpi=300)

## check correlation with posterior

cor_A_nonpretend <- cor.test(E2.half_game_first_clicks %>%
                               filter(test_part=='nonpretend_half_game' & grid_number=='A') %>% 
                               pull(n),scan(text= posteriorA, what = numeric(), sep="," , quiet = TRUE));

cor_A_pretend <- cor.test(E2.half_game_first_clicks %>%
                               filter(test_part=='pretend_half_game' & grid_number=='A') %>% 
                               pull(n),scan(text= posteriorA, what = numeric(), sep="," , quiet = TRUE));

cor_A_pretend_nonpretend <- cor.test(E2.half_game_first_clicks %>%
             filter(test_part=='nonpretend_half_game' & grid_number=='A') %>% 
             pull(n), 
             E2.half_game_first_clicks %>%
             filter(test_part=='pretend_half_game' & grid_number=='A') %>% 
             pull(n))

cor_B_nonpretend <- cor.test(E2.half_game_first_clicks %>%
                               filter(test_part=='nonpretend_half_game' & grid_number=='B') %>% 
                               pull(n),scan(text= posteriorB, what = numeric(), sep="," , quiet = TRUE));

cor_B_pretend <- cor.test(E2.half_game_first_clicks %>%
                               filter(test_part=='pretend_half_game' & grid_number=='B') %>% 
                               pull(n),scan(text= posteriorB, what = numeric(), sep="," , quiet = TRUE));

cor_B_pretend_nonpretend <- cor.test(E2.half_game_first_clicks %>%
             filter(test_part=='nonpretend_half_game' & grid_number=='B') %>% 
             pull(n), 
             E2.half_game_first_clicks %>%
             filter(test_part=='pretend_half_game' & grid_number=='B') %>% 
             pull(n))

```