### 2023-04-15 analysis/formatDataForTuring.Rmd
orthogonalize

---
title: "Format Datan For Turing"
author: "Matan Mazor"
date: "2023-04-07"
output: html_document
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)

# Load packages with groundhog
library(groundhog)
groundhog.library(c(
  'png',
  'grid',
  'ggplot2',
  'svglite',
  'xtable',
  'papaja',
  'tidyverse',
  'broom',
  'cowplot',
  'MESS', # for AUCs
  'lsr', # for effect sizes
  'pwr', # for power calculations
  'brms', # for mixed effects modeling
  'BayesFactor', # for Bayesian t test
  'jsonlite', # parsing data from sort_trial
  'caret', #for cross validation
  'ggrepel', #for word scatterplots
  'pracma' # for GramSchmidt
), '2022-12-01')

# Load workspace (after running preregisteredMethodsAndResults.Rmd, 
# exploratoryResults.Rmd, and all the scripts that are linked to from these 
# documents). 
load('../.RData')
```
## Calculate information gain

```{r hangman-information-gain, echo=FALSE}

get_word_state <- function(word, asked) {
  word_state = ''
  for (letter in strsplit(word,'')[[1]]) {
    value <- ifelse(grepl(toupper(letter), paste(toupper(asked),' ', sep='')), letter, '_')
    word_state <- paste(word_state, value, sep='')
  }
  return(toupper(word_state))
}

word_is_consistent <- function(word_state, asked,  word) {
  return(get_word_state(word,asked)==toupper(word_state))
}

get_word_entropy <- function(word_probabilities) {

    entropy <- -sum(unlist(word_probabilities$posterior)*
                    (unlist(log2(word_probabilities$posterior))),na.rm=TRUE);
  return(entropy)
};

get_letter_eig <- function(word_state, asked, category, 
                                          entropy, word_probabilities, letter) {

  if (grepl(toupper(letter),toupper(asked))) {
    return(NA)
  } else {
    word_probabilities <- word_probabilities %>%
      rowwise() %>%
      mutate(new_state = get_word_state(word,paste(asked,letter,sep=''))) %>%
      group_by(new_state) %>%
      summarize(prob = sum(posterior),
                entropy = -sum(unlist(posterior)*
                    (unlist(log2(posterior/sum(posterior)))),na.rm=TRUE))
    expected_entropy = sum(unlist(word_probabilities$prob)*unlist(word_probabilities$entropy))
    return(entropy-expected_entropy)
  }
}

#expected information gain
get_eig <- function(word_state,asked,category) {
  print(word_state)
  word_probabilities = get_word_probabilities(word_state,asked,category);
  entropy = get_word_entropy(word_probabilities)
  #if entropy is 0 there is no more uncertainty to resolve
  if (entropy==0) {
    print('entropy is 0')
    return(paste(rep(0,26),collapse=','))
  } else {
    letter_eig = c();
    for (letter in toupper(letters)) {
      if (grepl(letter,asked)) {
        letter_eig = c(letter_eig,
                                 NA)
      } else {
        letter_eig = c(letter_eig,
                               get_letter_eig(word_state,asked,category,entropy, 
                                              word_probabilities, letter)
      )
      }
    }
  }
  return(paste(letter_eig,collapse=','))
}

# E4.click_log_with_information_gain <- E4.click_log %>%
#   dplyr::select(subj_id,genuine_first,test_part,word,category,word_state,asked,click_number,letter,RT,hit_bin) %>%
#   rowwise()%>%
#   mutate(eig = get_eig(word_state,asked,category))

load('E4eig.RData')

# get_letter_posterior <- function(word_state, asked, category) {
# 
#   word_probabilities <- get_word_probabilities(word_state, asked, category);
# 
#   letter_probabilities = c();
#   for (letter in toupper(letters)) {
#     if (grepl(letter,asked)) {
#       letter_probabilities = c(letter_probabilities,
#                                NA)
#     } else {
#       letter_probabilities = c(letter_probabilities,
#                              word_probabilities %>%
#                                filter(grepl(letter,toupper(word))) %>%
#                                pull(posterior) %>%
#                                sum()
#     )
#     }
#   }
#   posterior = letter_probabilities/sum(letter_probabilities, na.rm=T)
# 
#   return(paste(posterior,collapse=','))
# 
# };
# 
# get_letter_prob_posterior <- function(asked) {
# 
#   letter_probabilities <- letter_df %>%
#     rowwise() %>%
#     mutate(prob = ifelse(
#       grepl(toupper(letter),toupper(asked)),
#       NA,
#       frequency)
#       ) %>%
#     ungroup()%>%
#     mutate(
#       posterior = prob/sum(prob,na.rm=T)
#     )
# 
#   return(paste(letter_probabilities$posterior,collapse=','))
# 
# };
# 
# get_p_letter <- function(posterior, letter) {
#   posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
#   return(posterior[which(toupper(letters)==letter)])
# }
# 
# get_p_letter_rank <- function(posterior, letter) {
#   posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
#   rank_posterior = rank(-posterior)
#   return(rank_posterior[which(toupper(letters)==letter)])
# }
# 
# get_letter_posterior_entropy <- function(posterior) {
#   posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
#   posterior[which(posterior==0)]=NA; #to avoid infinity*0
#   entropy <- -sum(unlist(posterior)*(unlist(log(posterior))),na.rm=TRUE);
#   return(entropy);
# }

# E4.click_log_with_boards <- E4.click_log %>%
#   dplyr::select(subj_id,genuine_first,test_part,word,category,word_state,asked,click_number,letter,RT,hit_bin) %>%
#   rowwise()%>%
#   mutate(posterior = get_letter_posterior(word_state,asked,category),
#          p_click = get_p_letter(posterior,letter),
#          entropy = get_letter_posterior_entropy(posterior),
#          p_click_rank = get_p_letter_rank(posterior,letter));

# get_random_letter <- function(asked) {
# 
#   asked = strsplit(tolower(asked),'')[[1]];
#   not_asked = setdiff(letters,asked)
#   return(toupper(not_asked)%>%sample(1))
# 
# }
# 
# get_most_frequent_letter <- function(asked) {
# 
#    lp <- letter_df %>%
#     rowwise() %>%
#     mutate(prob = ifelse(
#       grepl(toupper(letter),toupper(asked)),
#       NA,
#       frequency)
#       ) %>%
#     ungroup()
# 
#      return(lp[which.max(lp$prob),]$letter%>%toupper())
# 
# }

# E4.click_log_with_boards_random <- E4.click_log_with_boards %>%
#   mutate(letter = get_random_letter(asked),
#          p_click = get_p_letter(posterior,letter),
#          p_click_rank = get_p_letter_rank(posterior,letter),
#          hit_bin = tolower(letter) %in% strsplit(tolower(word),'')[[1]],
#          test_part='random');

# E4.click_log_with_boards_most_frequent <- E4.click_log_with_boards %>%
#   mutate(letter = get_most_frequent_letter(asked),
#          p_click = get_p_letter(posterior,letter),
#          p_click_rank = get_p_letter_rank(posterior,letter),
#          hit_bin = tolower(letter) %in% strsplit(tolower(word),'')[[1]],
#          test_part='most_frequent');

# E4.letter_prob_click_log_with_boards <- E4.click_log %>%
#   dplyr::select(subj_id,genuine_first,test_part,word,category,word_state,asked,click_number,letter,RT,hit_bin) %>%
#   rowwise()%>%
#   mutate(posterior = get_letter_prob_posterior(asked),
#          p_click = get_p_letter(posterior,letter),
#          entropy = get_letter_posterior_entropy(posterior),
#          p_click_rank = get_p_letter_rank(posterior,letter));

# E4.letter_prob_click_log_with_boards_random <- E4.letter_prob_click_log_with_boards %>%
#   mutate(letter = get_random_letter(asked),
#          p_click = get_p_letter(posterior,letter),
#          p_click_rank = get_p_letter_rank(posterior,letter),
#          hit_bin = tolower(letter) %in% strsplit(tolower(word),'')[[1]],
#          test_part='random');

# save(E4.click_log_with_boards,
#      E4.click_log_with_boards_random,
#      E4.click_log_with_boards_most_frequent,
#      E4.letter_prob_click_log_with_boards,
#      E4.letter_prob_click_log_with_boards_random,
#      file='E4.click_log_with_boards.RData')

load('../docs/E4.click_log_with_boards.RData') # this file includes all the above dataframes.

```

```{r format}

E4.letters_posterior_wide <- E4.click_log_with_boards %>% 
  dplyr::select(subj_id, test_part,word,posterior,letter, hit_bin) %>%
  separate(posterior, sep=",", into=paste0("post_",LETTERS)) %>%
  mutate_at(vars(post_A:post_Z), as.numeric) %>% 
  mutate(across(starts_with("post"), ~ifelse(is.na(.), 0, 1), .names = "{sub('post', 'mask', .col)}"),
         across(starts_with("post"), ~ifelse(is.na(.), 0, .)))

E4.letters_prior_wide <- E4.letter_prob_click_log_with_boards %>% 
  dplyr::select(subj_id, test_part,word,posterior,letter) %>%
  separate(posterior, sep=",", into=paste0("prior_",LETTERS)) %>%
  mutate_at(vars(prior_A:prior_Z), as.numeric) %>% 
  mutate(across(starts_with("prior"), ~ifelse(is.na(.), 0, 1), .names = "{sub('prior', 'mask', .col)}"),
         across(starts_with("prior"), ~ifelse(is.na(.), 0, .)))

E4.letters_eig_wide <- E4.click_log_with_information_gain %>% 
  dplyr::select(subj_id, test_part,word,eig,letter) %>%
  separate(eig, sep=",", into=paste0("eig_",LETTERS)) %>%
  mutate_at(vars(eig_A:eig_Z), as.numeric) %>% 
  mutate(across(starts_with("eig"), ~ifelse(is.na(.), 0, .)))

E4.letters_features_wide <- E4.letters_posterior_wide %>%
  merge(E4.letters_prior_wide) %>%
  merge(E4.letters_eig_wide)

E4.letters_features_wide %>%
  write.csv('dataForTuring/E4.csv')
```

```{r normalize-features}
E4.letters_features_normalized <- E4.letters_features_wide;

post_columns <- paste0("post_",LETTERS)
prior_columns <- paste0("prior_",LETTERS)
eig_columns <- paste0("eig_",LETTERS)

for (subj in E4.letters_features_normalized$subj_id%>%unique()) {

  subj_rows <- which(E4.letters_features_normalized$subj_id==subj)
  post_subset <- E4.letters_features_normalized[subj_rows,post_columns];
  # Calculate the global mean and standard deviation for the subset
  post_mean <- mean(as.vector(unlist((post_subset))), na.rm = TRUE)
  post_sd <- sd(as.vector(unlist((post_subset))), na.rm = TRUE)

  # Normalize the selected columns
  normalized_post_subset <- post_subset %>% mutate(across(everything(), ~ (. - post_mean) / post_sd))

  # Replace the original columns with the normalized ones
  E4.letters_features_normalized[subj_rows, post_columns] <- normalized_post_subset

  prior_subset <- E4.letters_features_normalized[subj_rows,prior_columns];
  # Calculate the global mean and standard deviation for the subset
  prior_mean <- mean(as.vector(unlist((prior_subset))), na.rm = TRUE)
  prior_sd <- sd(as.vector(unlist((prior_subset))), na.rm = TRUE)

  # Normalize the selected columns
  normalized_prior_subset <- prior_subset %>% mutate(across(everything(), ~ (. - prior_mean) / prior_sd))

  # Replace the original columns with the normalized ones
  E4.letters_features_normalized[subj_rows, prior_columns] <- normalized_prior_subset

  eig_subset <- E4.letters_features_normalized[subj_rows,eig_columns];
  # Calculate the global mean and standard deviation for the subset
  eig_mean <- mean(as.vector(unlist((eig_subset))), na.rm = TRUE)
  eig_sd <- sd(as.vector(unlist((eig_subset))), na.rm = TRUE)

  # Normalize the selected columns
  normalized_eig_subset <- eig_subset %>% mutate(across(everything(), ~ (. - eig_mean) / eig_sd))

  # Replace the original columns with the normalized ones
  E4.letters_features_normalized[subj_rows, eig_columns] <- normalized_eig_subset;
}

E4.letters_features_normalized %>%
  write.csv('dataForTuring/E4norm.csv')

```

```{r orthogonalize-eig}
E4.letters_features_orth <- E4.letters_features_normalized;

post_columns <- paste0("post_",LETTERS)
prior_columns <- paste0("prior_",LETTERS)
eig_columns <- paste0("eig_",LETTERS)
mask_columns <- paste0("mask_", LETTERS)

num_rows <- nrow(E4.letters_features_orth)

pb = txtProgressBar(min = 0, max = num_rows, initial = 0)

for (i_row in 1:num_rows) {

  mask_subset <- unlist(E4.letters_features_orth[i_row,mask_columns], use.names=F);
  mask_indices <- which(mask_subset==1);
  post_subset <- unlist(E4.letters_features_orth[i_row,post_columns[mask_indices]], use.names=F);
  prior_subset <- unlist(E4.letters_features_orth[i_row,prior_columns[mask_indices]], use.names=F);
  eig_subset <- unlist(E4.letters_features_orth[i_row,eig_columns[mask_indices]], use.names=F);

    matrix <- matrix(c(post_subset,prior_subset,eig_subset),length(mask_indices),3)
    if (Rank(mask_indices)==3) {
    orthogonalized_matrix <- gramSchmidt(matrix)$Q

    orth_post_subset <- orthogonalized_matrix[,1];
    orth_prior_subset <- orthogonalized_matrix[,2];
    orth_eig_subset <- orthogonalized_matrix[,3];
  } else {
    orth_post_subset <- post_subset;
    orth_prior_subset <- prior_subset;
    orth_eig_subset <- eig_subset;
  }

 # Replace the original columns with the orthogonalized ones
  E4.letters_features_orth[i_row, post_columns[mask_indices]] <- orth_post_subset
  E4.letters_features_orth[i_row, prior_columns[mask_indices]] <- orth_prior_subset
  E4.letters_features_orth[i_row, eig_columns[mask_indices]] <- orth_eig_subset

  setTxtProgressBar(pb, i_row)

}

close(pb)

# NORMALIZE!

num_subj <- E4.letters_features_orth$subj_id%>%unique()%>%length()
pb = txtProgressBar(min = 0, max = num_subj, initial = 0)

i_subj=0
for (subj in E4.letters_features_orth$subj_id%>%unique()) {
  i_subj = i_subj+1
  subj_rows <- which(E4.letters_features_orth$subj_id==subj)
  post_subset <- E4.letters_features_orth[subj_rows,post_columns];
  # Calculate the global mean and standard deviation for the subset
  post_mean <- mean(as.vector(unlist((post_subset))), na.rm = TRUE)
  post_sd <- sd(as.vector(unlist((post_subset))), na.rm = TRUE)

  # Normalize the selected columns
  normalized_post_subset <- post_subset %>% mutate(across(everything(), ~ (. - post_mean) / post_sd))

  # Replace the original columns with the normalized ones
  E4.letters_features_orth[subj_rows, post_columns] <- normalized_post_subset

  prior_subset <- E4.letters_features_orth[subj_rows,prior_columns];
  # Calculate the global mean and standard deviation for the subset
  prior_mean <- mean(as.vector(unlist((prior_subset))), na.rm = TRUE)
  prior_sd <- sd(as.vector(unlist((prior_subset))), na.rm = TRUE)

  # Normalize the selected columns
  normalized_prior_subset <- prior_subset %>% mutate(across(everything(), ~ (. - prior_mean) / prior_sd))

  # Replace the original columns with the normalized ones
  E4.letters_features_orth[subj_rows, prior_columns] <- normalized_prior_subset

  eig_subset <- E4.letters_features_orth[subj_rows,eig_columns];
  # Calculate the global mean and standard deviation for the subset
  eig_mean <- mean(as.vector(unlist((eig_subset))), na.rm = TRUE)
  eig_sd <- sd(as.vector(unlist((eig_subset))), na.rm = TRUE)

  # Normalize the selected columns
  normalized_eig_subset <- eig_subset %>% mutate(across(everything(), ~ (. - eig_mean) / eig_sd))

  # Replace the original columns with the normalized ones
  E4.letters_features_orth[subj_rows, eig_columns] <- normalized_eig_subset;

  setTxtProgressBar(pb, i_subj)

}

E4.letters_features_orth %>%
  write.csv('dataForTuring/E4orth.csv')

```

```{r}

E4.letters_features_wide_misses <- E4.letters_features_wide %>%
  filter(!hit_bin)

E4.subjects_with_enough_misses <- E4.letters_features_wide_misses %>%
  group_by(subj_id,test_part) %>%
  summarise(n=n()) %>%
  group_by(subj_id) %>%
  summarise(n=min(n)) %>%
  filter(n>5) %>%
  pull(subj_id)

E4.letters_features_wide_misses %>%
  filter(subj_id %in% E4.subjects_with_enough_misses) %>%
  write.csv('dataForTuring/E4misses.csv')

```
## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.