### 2023-05-16 analysis/genderDifferences.Rmd
gender differences

---
title: "GenderDifferences"
author: "Matan Mazor"
date: "2023-05-16"
output: html_document
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)

# Load packages with groundhog
library(groundhog)
groundhog.library(c(
  'png',
  'grid',
  'ggplot2',
  'svglite',
  'xtable',
  'papaja',
  'tidyverse',
  'broom',
  'cowplot',
  'MESS', # for AUCs
  'lsr', # for effect sizes
  'pwr', # for power calculations
  'brms', # for mixed effects modeling
  'BayesFactor', # for Bayesian t test
  'jsonlite', # parsing data from sort_trial
  'caret', #for cross validation
  'ggrepel', #for word scatterplots
  'memoise' # for function memoization
), '2022-12-01')

# Load workspace (after running preregisteredMethodsAndResults.Rmd, 
# exploratoryResults.Rmd, and all the scripts that are linked to from these 
# documents). 
load('../.RData')
```

## Load model coefficients

```{r create-gender-dfs, echo=FALSE}

E2.gender <- E2.export %>%
  filter(status=='APPROVED') %>%
  merge(E2.df %>%
          mutate(participant_id=PROLIFIC_PID) %>%
          dplyr::select(participant_number,participant_id,subj_id),by='participant_id') %>%
  dplyr::select(subj_id,Sex) %>%
  group_by(subj_id) %>%
  summarise(sex=Sex[1]) %>%
  rowwise()

E4.gender <- E4.export %>%
  filter(status=='APPROVED') %>%
  merge(E4.df %>%
          mutate(participant_id=PROLIFIC_PID) %>%
          dplyr::select(participant_number,participant_id,subj_id),by='participant_id') %>%
  dplyr::select(subj_id,Sex) %>%
  group_by(subj_id) %>%
  summarise(sex=Sex[1]) %>%
  rowwise()
```

```{r p_click, echo-FALSE, cache=TRUE}

E2.p_click_by_sex <- E2.p_click_summary %>%
  merge(E2.gender) %>%
  filter(sex %in% c('Female','Male'))

E2.p_click_by_sex_summary <- E2.p_click_by_sex %>% 
  group_by(test_part,sex) %>%
  summarise(p_click=mean(p_click_rank))

E4.p_click_by_sex <- E4.p_click_summary %>%
  merge(E4.gender) %>%
  filter(sex %in% c('Female','Male'))

E4.p_click_by_sex_summary <- E4.p_click_by_sex %>% 
  group_by(test_part,sex) %>%
  summarise(p_click=mean(p_click_rank))

apa_table(
  E2.p_click_by_sex_summary
  , caption = "p_click by gender in Battleship."
  , escape = TRUE
)

apa_table(
  E4.p_click_by_sex_summary
  , caption = "p_click by gender in Hangman."
  , escape = TRUE
)
```

We find no differences between the two sexes in the posterior probability of a hit, in neither pretend nor non-pretend games.

```{r hangman-halfgames, echo=F, cache=T}

E4.first_guess_by_sex <- E4.first_click_half_games %>%
  merge(E4.gender) %>%
  filter(sex %in% c('Female','Male'))

E4.papaya_by_sex <- E4.first_guess_by_sex %>%
  filter(word=='papaya') %>%
  group_by(test_part,sex) %>%
  summarise(banana=mean(bananahand & makes_sense),
            papaya = mean(!bananahand & makes_sense),
            other = mean(!makes_sense))

E4.hair_by_sex <- E4.first_guess_by_sex %>%
  filter(word=='hair') %>%
  group_by(test_part,sex) %>%
  summarise(hand=mean(bananahand & makes_sense),
            hair = mean(!bananahand & makes_sense),
            other = mean(!makes_sense))

apa_table(
  E4.papaya_by_sex
  , caption = "first click by gender in Hangman half games, when the target word is papaya. Showing proportion of guesses that are consistent with banana, papaya, or other."
  , escape = TRUE
)

apa_table(
  E4.hair_by_sex
  , caption = "first click by gender in Hangman half games, when the target word is hair. Showing proportion of guesses that are consistent with hand, hair, or other."
  , escape = TRUE
)

```