### 2024-08-22 analysis/JudgmentBasedAnalysis.Rmd
judgment based analysis

---
title: "Judgment Based Analysis"
output: html_document
date: "2024-08-22"
---

```{r analysis-preferences, warning=F, message=F, echo=FALSE}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, warning=F, message=F)

# Load packages with groundhog
library(groundhog)
groundhog.library(c(
  'png',
  'grid',
  'ggplot2',
  'svglite',
  'xtable',
  'papaja',
  'tidyverse',
  'broom',
  'cowplot',
  'MESS', # for AUCs
  'lsr', # for effect sizes
  'pwr', # for power calculations
  'brms', # for mixed effects modeling
  'BayesFactor', # for Bayesian t test
  'jsonlite', # parsing data from sort_trial
  'caret', #for cross validation
  'ggrepel' #for word scatterplots
), '2023-12-01')

# Load workspace (after running preregisteredMethodsAndResults.Rmd, 
# exploratoryResults.Rmd, and all the scripts that are linked to from these 
# documents). 
load('../.RData')

papaja::r_refs("r-references.bib")

#color scheme
colors <-  list()
colors$p <- "#404080"
colors$np <- "#69b3a2"
colors$random <- 'gray'
colors$greedy <- 'black'
```

## Construct a subj_id mapping

```{r subj_id_map, warning=F, message=F, echo=FALSE}

E2.subj_mapping <- E2.df %>% 
  group_by(subj_id) %>%
  summarise(jatos_id=subject_identifier[1])

E4.subj_mapping <- E4.df %>% 
  group_by(subj_id) %>%
  summarise(jatos_id=subject_identifier[1])
```

## Organize games according to judgments

```{r games_by_judgments, warning=F, message=F, echo=FALSE}

# Transform the dataframe
E2.judgment <- E2.judge_df %>%
  # Create rows for cheat_player
  mutate(player = cheat_player, judged_pretend = correct, judge=subj_id) %>%
  select(player, grid_number, judged_pretend, judge) %>%
  # Bind rows for genuine_player
  bind_rows(
    E2.judge_df %>%
      mutate(player = genuine_player, judged_pretend = 1 - correct, judge=subj_id) %>%
      select(player, grid_number, judged_pretend, judge)
  ) %>%
  arrange(grid_number, player, judged_pretend)%>%
  rename(jatos_id=player) %>%
  merge(E2.subj_mapping, by='jatos_id')

E4.judgment <- E4.judge_df %>%
  mutate(judged_pretend = ifelse(correct_response=='P',correct,1-correct)) %>%
  select(player,word,judged_pretend, subj_id)%>%
  rename(jatos_id=player,
         judge=subj_id) %>%
  merge(E4.subj_mapping, by='jatos_id')
```

## Merge with main dfs

```{r merge_with_df, warning=F, message=F, echo=FALSE}

E2.click_df_by_judgment <- E2.click_df %>% 
  merge(E2.judgment, by=c('subj_id','grid_number')) %>%
  mutate(judged = ifelse(judged_pretend==1,'pretend','nonpretend'))

E4.click_df_by_judgment <- E4.click_df %>% 
  merge(E4.judgment, by=c('subj_id','word')) %>%
  mutate(judged = ifelse(judged_pretend==1,'pretend','nonpretend'))

E2.click_log_with_boards_by_judgment <- E2.click_log_with_boards %>% 
  merge(E2.judgment, by=c('subj_id','grid_number')) %>%
  mutate(judged = ifelse(judged_pretend==1,'pretend','nonpretend'))

E4.click_log_with_boards_by_judgment <- E4.click_log_with_boards %>% 
  merge(E4.judgment, by=c('subj_id','word')) %>%
  mutate(judged = ifelse(judged_pretend==1,'pretend','nonpretend'))

```

## Number of misses

```{r num_misses, warning=F, message=F, echo=FALSE}

E2.median_clicks_by_judgment  <- E2.click_df_by_judgment  %>%
  group_by(judge,judged) %>%
  summarise(num_clicks=median(num_clicks)) %>%
  spread(judged,num_clicks)%>%
  mutate(diff=pretend-nonpretend);

p<- E2.click_df_by_judgment %>%
    group_by(judged,judge)%>%
    summarise(num_clicks=median(num_clicks))%>%
    ggplot(aes(x=num_clicks,fill=judged))+
    geom_bar( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    labs(x='median number of clicks', y='number of judges')+
    theme(legend.position=c(0.10,0.85)) +
    theme_classic()

ggsave('../docs/figures/judgment/E2_num_clicks.png',p,width=5,height=2, dpi=300)

E4.median_clicks_by_judgment  <- E4.click_df_by_judgment  %>%
  group_by(judge,judged) %>%
  summarise(num_clicks=median(num_clicks)) %>%
  spread(judged,num_clicks) %>%
  mutate(diff=pretend-nonpretend);

p<- E4.click_df_by_judgment %>%
    group_by(judged,judge)%>%
    summarise(num_clicks=median(num_clicks))%>%
    ggplot(aes(x=num_clicks,fill=judged))+
    geom_bar( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    labs(x='median number of clicks', y='number of judges')+
    theme(legend.position=c(0.10,0.85)) +
    theme_classic()

ggsave('../docs/figures/judgment/judgment/E2_num_clicks.png',p,width=5,height=2, dpi=300)
```

Games with fewer misses are more likely to be categorised as pretend games in Battleship (`r apa_print(t.test(E2.median_clicks_by_judgment%>%pull(diff)))$full_result`) but not in Hangman (`r apa_print(t.test(E4.median_clicks_by_judgment%>%pull(diff)))$full_result`).

```{r merge_with_df, warning=F, message=F, echo=FALSE}

E2.median_clicks_by_judgment  <- E2.click_df_by_judgment  %>%
  group_by(subj_id,judged) %>%
  summarise(num_clicks=median(num_clicks)) %>%
  spread(judged,num_clicks);

E2.lucky_pretenders <- E2.click_df%>%filter(num_clicks==7)%>%pull(subj_id)%>%unique();

E2.median_clicks_by_judgment_filtered <- E2.click_df_by_judgment %>%
  filter(!(subj_id %in% E2.lucky_pretenders))%>%
  group_by(subj_id,judged, genuine_first) %>%
  summarise(num_clicks=median(num_clicks)) %>%
  spread(judged,num_clicks);

p<- E2.click_df_by_judgment %>%
    ggplot(aes(x=num_clicks,fill=judged))+
    geom_bar( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    labs(x='number of clicks', y='number of games')+
    theme(legend.position=c(0.10,0.85)) +
    theme_classic()

ggsave('../docs/figures/judgment/E2_num_clicks.png',p,width=5,height=2, dpi=300)

E4.median_clicks_by_judgment  <- E4.click_df_by_judgment  %>%
  group_by(subj_id,judged) %>%
  summarise(num_clicks=median(num_clicks)) %>%
  spread(judged,num_clicks);

p<- E4.click_df_by_judgment %>%
    ggplot(aes(x=num_clicks,fill=judged))+
    geom_bar( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    labs(x='number of clicks', y='number of games')+
    theme(legend.position=c(0.10,0.85)) +
    theme_classic()

ggsave('../docs/figures/judgment/judgment/E2_num_clicks.png',p,width=5,height=2, dpi=300)
```