### 2024-08-23 analysis/JudgmentBasedAnalysis.Rmd
RT analysis, accuracy analysis, guesses by word

---
title: "Judgment Based Analysis"
output: html_document
date: "2024-08-22"
---

```{r analysis-preferences, warning=F, message=F, echo=FALSE}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, warning=F, message=F)

# Load packages with groundhog
library(groundhog)
groundhog.library(c(
  'png',
  'grid',
  'ggplot2',
  'svglite',
  'xtable',
  'papaja',
  'tidyverse',
  'broom',
  'cowplot',
  'MESS', # for AUCs
  'lsr', # for effect sizes
  'pwr', # for power calculations
  'brms', # for mixed effects modeling
  'BayesFactor', # for Bayesian t test
  'jsonlite', # parsing data from sort_trial
  'caret', #for cross validation
  'ggrepel' #for word scatterplots
), '2023-12-01')

# Load workspace (after running preregisteredMethodsAndResults.Rmd, 
# exploratoryResults.Rmd, and all the scripts that are linked to from these 
# documents). 
load('../.RData')

papaja::r_refs("r-references.bib")

#color scheme
colors <-  list()
colors$p <- "#404080"
colors$np <- "#69b3a2"
colors$random <- 'gray'
colors$greedy <- 'black'
```

## Construct a subj_id mapping

```{r subj_id_map, warning=F, message=F, echo=FALSE}

E2.subj_mapping <- E2.df %>% 
  group_by(subj_id) %>%
  summarise(jatos_id=subject_identifier[1])

E4.subj_mapping <- E4.df %>% 
  group_by(subj_id) %>%
  summarise(jatos_id=subject_identifier[1])
```

## Organize games according to judgments

```{r games_by_judgments, warning=F, message=F, echo=FALSE}

# Transform the dataframe
E2.judgment <- E2.judge_df %>%
  # Create rows for cheat_player
  mutate(player = cheat_player, judged_pretend = correct, judge_id=subj_id) %>%
  select(player, grid_number, judged_pretend, judge_id) %>%
  # Bind rows for genuine_player
  bind_rows(
    E2.judge_df %>%
      mutate(player = genuine_player, judged_pretend = 1 - correct, judge_id=subj_id) %>%
      select(player, grid_number, judged_pretend, judge_id)
  ) %>%
  arrange(grid_number, player, judged_pretend)%>%
  rename(jatos_id=player) %>%
  merge(E2.subj_mapping, by='jatos_id')

E4.judgment <- E4.judge_df %>%
  mutate(judged_pretend = ifelse(correct_response=='P',correct,1-correct)) %>%
  select(player,word,judged_pretend, subj_id)%>%
  rename(jatos_id=player,
         judge_id=subj_id) %>%
  merge(E4.subj_mapping, by='jatos_id')
```

## Merge with main dfs

```{r merge_with_df, warning=F, message=F, echo=FALSE}

E2.click_df_by_judgment <- E2.click_df %>% 
  merge(E2.judgment, by=c('subj_id','grid_number')) %>%
  mutate(judged = ifelse(judged_pretend==1,'pretend','nonpretend'))

E4.click_df_by_judgment <- E4.click_df %>% 
  merge(E4.judgment, by=c('subj_id','word')) %>%
  mutate(judged = ifelse(judged_pretend==1,'pretend','nonpretend'))

E2.click_log_with_boards_by_judgment <- E2.click_log_with_boards %>% 
  merge(E2.judgment, by=c('subj_id','grid_number')) %>%
  mutate(judged = ifelse(judged_pretend==1,'pretend','nonpretend'))

E4.click_log_with_boards_by_judgment <- E4.click_log_with_boards %>% 
  merge(E4.judgment, by=c('subj_id','word')) %>%
  mutate(judged = ifelse(judged_pretend==1,'pretend','nonpretend'))

E2.click_log_by_judgment <- E2.click_log %>% 
  merge(E2.judgment, by=c('subj_id','grid_number')) %>%
  mutate(judged = ifelse(judged_pretend==1,'pretend','nonpretend'))

E4.click_log_by_judgment <- E4.click_log %>% 
  merge(E4.judgment, by=c('subj_id','word')) %>%
  mutate(judged = ifelse(judged_pretend==1,'pretend','nonpretend'))

```

## Number of misses

```{r num_misses, warning=F, message=F, echo=FALSE}

E2.median_clicks_by_judgment  <- E2.click_df_by_judgment  %>%
  group_by(judge_id,judged) %>%
  summarise(num_clicks=median(num_clicks)) %>%
  spread(judged,num_clicks)%>%
  mutate(diff=pretend-nonpretend);

p<- E2.click_df_by_judgment %>%
    group_by(judged,judge_id)%>%
    summarise(num_clicks=median(num_clicks))%>%
    ggplot(aes(x=num_clicks,fill=judged))+
    geom_bar( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    labs(x='median number of clicks', y='number of judges')+
    theme(legend.position=c(0.10,0.85)) +
    theme_classic()

ggsave('../docs/figures/judgment/E2_num_clicks.png',p,width=5,height=2, dpi=300)

E4.median_clicks_by_judgment  <- E4.click_df_by_judgment  %>%
  group_by(judge_id,judged) %>%
  summarise(num_clicks=median(num_clicks)) %>%
  spread(judged,num_clicks) %>%
  mutate(diff=pretend-nonpretend);

p<- E4.click_df_by_judgment %>%
    group_by(judged,judge_id)%>%
    summarise(num_clicks=median(num_clicks))%>%
    ggplot(aes(x=num_clicks,fill=judged))+
    geom_bar( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    labs(x='median number of clicks', y='number of judges')+
    theme(legend.position=c(0.10,0.85)) +
    theme_classic()

ggsave('../docs/figures/judgment/judgment/E2_num_clicks.png',p,width=5,height=2, dpi=300)
```
## Number of misses, controlling for word

```{r num_misses, warning=F, message=F, echo=FALSE}

E4.median_clicks_per_word <- E4.click_df %>%
  group_by(word)%>%
  summarise(med_clicks=median(num_clicks))

E4.corrected_clicks_by_judgment  <- E4.click_df_by_judgment  %>%
  merge(E4.median_clicks_per_word,by='word')%>%
  mutate(relative_clicks=num_clicks-med_clicks)%>%
  group_by(judge_id,judged) %>%
  summarise(num_clicks=median(relative_clicks)) %>%
  spread(judged,num_clicks) %>%
  mutate(diff=pretend-nonpretend);

E4.corrected_clicks_by_judgment_mean  <- E4.click_df_by_judgment  %>%
  merge(E4.median_clicks_per_word,by='word')%>%
  mutate(relative_clicks=num_clicks-med_clicks)%>%
  group_by(judge_id,judged) %>%
  summarise(num_clicks=mean(relative_clicks)) %>%
  spread(judged,num_clicks) %>%
  mutate(diff=pretend-nonpretend);

```
No effect in Hangman even when controlling for average number of clicks per word (`r APA_print(E4.corrected_clicks_by_judgment$diff%>%t.test())$statistic`), but a weak effect in the opposite direction (such that longer games are judged to be pretend games) when using mean instead of median (which might actually make sense; `r APA_print(E4.corrected_clicks_by_judgment$diff%>%t.test())$statistic`).

## Misses per word

```{r num_misses, warning=F, message=F, echo=FALSE}

E4.clicks_by_word_by_judgment  <- E4.click_df_by_judgment  %>%
  group_by(word,judged) %>%
  summarise(num_clicks=mean(num_clicks)) %>%
  spread(judged,num_clicks) %>%
  mutate(diff=pretend-nonpretend);

E4.clicks_by_word_by_judgment %>%
  ggplot(aes(x=nonpretend,y=pretend, label=word))+
  geom_label()+
  geom_abline(slope=1,intercept=0)+
  labs(x='judged non-pretend',y='judged pretend')
```

The same effect to what we see when categorising based on true pretend/non-pretend, such that people are more likely to judge unusually short or long games as nonpretend games.

## Game duration

```{r duration, warning=F, message=F, echo=FALSE}

E2.duration_by_judgment <- E2.click_log_with_boards_by_judgment %>% 
  group_by(judge_id,judged,grid_number) %>%
  summarise(duration=sum(RT)) %>% 
  group_by(judge_id,judged) %>% 
  summarise(duration=median(duration)) %>% 
  spread(judged,duration) %>%
  mutate(diff=pretend-nonpretend)

E4.duration_by_judgment <- E4.click_log_with_boards_by_judgment %>% 
  group_by(judge_id,judged,word) %>%
  summarise(duration=sum(RT)) %>% 
  group_by(judge_id,judged) %>% 
  summarise(duration=median(duration)) %>% 
  spread(judged,duration) %>%
  mutate(diff=pretend-nonpretend)

```
Long and short games were similarly likely to be categorised as pretend games in Battleship (`r apa_print(t.test(E2.duration_by_judgment%>%pull(diff)))$full_result`) and in Hangman (`r apa_print(t.test(E4.duration_by_judgment%>%pull(diff)))$full_result`).

```{r RT_by_accuracy}

N_perm <- 1000;
bootstrap_error <- function(x, N_perm) {
  N <- length(x)
  medians = c();
  for (i in 1:N_perm) {
    medians = c(medians,sample(x,replace=TRUE,size=N)%>%median(., na.rm=T))
  };
  return(sd(medians))
}

E2.RT_by_lag1_by_judgment <- E2.click_log_by_judgment %>%
  filter(click_number>1) %>%
  group_by(judge_id,judged, lag1)%>%
  summarise(RT=median(RT))  %>%
  spread(lag1,RT,sep='_') %>%
  mutate(lag1_diff = lag1_TRUE-lag1_FALSE)

E2.RT_by_lag1_by_judgment_wide <- E2.RT_by_lag1_by_judgment %>%
  dplyr::select(judge_id,judged,lag1_diff) %>%
  spread(judged,lag1_diff)%>%
  mutate(diff=pretend-nonpretend)

E2.RT_by_hit_by_judgment <- E2.click_log_by_judgment %>%
  filter(click_number>0) %>%
  group_by(judge_id,judged,hit_bin)%>%
  summarise(RT=median(RT))  %>%
  spread(hit_bin,RT,sep='_') %>%
  mutate(hit_diff = hit_bin_TRUE-hit_bin_FALSE)

E2.RT_by_hit_by_judgment_wide <- E2.RT_by_hit_by_judgment %>%
  dplyr::select(judge_id,judged,hit_diff) %>%
  spread(judged,hit_diff)%>%
  mutate(diff=pretend-nonpretend)

E2.RT_by_lead1_by_judgment <- E2.click_log_by_judgment %>%
  group_by(judge_id,judged, lead1)%>%
  summarise(RT=median(RT))  %>%
  spread(lead1,RT,sep='_') %>%
  mutate(lead1_diff = lead1_TRUE-lead1_FALSE);

E2.RT_by_lead1_by_judgment_wide <- E2.RT_by_lead1_by_judgment %>%
  dplyr::select(judge_id,judged,lead1_diff) %>%
  spread(judged,lead1_diff)%>%
  mutate(diff=pretend-nonpretend)

E2.RT_by_hit_lags_by_judgment <-E2.RT_by_lag1_by_judgment %>%
  merge(E2.RT_by_hit_by_judgment) %>%
  merge(E2.RT_by_lead1_by_judgment)

E2.RT_by_hit_lags_long_by_judgment <- E2.RT_by_hit_lags_by_judgment %>%
  gather("position","difference",c("lag1_diff", "hit_diff", "lead1_diff"))%>%
  mutate(position = ifelse(position=="lag1_diff",1,
                                         ifelse(position=="hit_diff",0,-1))) %>%
  dplyr::select(judge_id,judged,position,difference);

p<- E2.RT_by_hit_lags_long_by_judgment %>% 
  filter(position<2 & position > -2)%>%
  group_by(judged,position) %>%
  summarise(median_difference=median(difference),
            sem_difference=bootstrap_error(difference,N_perm)) %>%
  ggplot(aes(x=position,color=judged,fill=judged,y=median_difference)) +
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  scale_color_manual(values=c("#69b3a2", "#404080")) +
  scale_x_continuous(breaks=-1:1,labels=c('hit on next trial', 'hit on current trial', 'hit on previous trial'))+
  geom_abline(slope=0,intercept=0,linetype=3)+
  geom_path(size=1)+
  geom_ribbon(aes(ymin=median_difference-sem_difference,ymax=median_difference+sem_difference),alpha=0.3,show.legend=F)+
  labs(fill=' ',x='', y='RT effect (ms) of hit on trial t')+
  theme_classic()

ggsave('figures/judgment/E2_RT_by_lag.png',p,width=4, height=3.5, dpi=300)

p

E4.RT_by_lag1_by_judgment <- E4.click_log_by_judgment %>%
  filter(click_number>1) %>%
  group_by(judge_id,judged, lag1)%>%
  summarise(RT=median(RT))  %>%
  spread(lag1,RT,sep='_') %>%
  mutate(lag1_diff = lag1_TRUE-lag1_FALSE)

E4.RT_by_lag1_by_judgment_wide <- E4.RT_by_lag1_by_judgment %>%
  dplyr::select(judge_id,judged,lag1_diff) %>%
  spread(judged,lag1_diff)%>%
  mutate(diff=pretend-nonpretend)

E4.RT_by_hit_by_judgment <- E4.click_log_by_judgment %>%
  filter(click_number>0) %>%
  group_by(judge_id,judged,hit_bin)%>%
  summarise(RT=median(RT))  %>%
  spread(hit_bin,RT,sep='_') %>%
  mutate(hit_diff = hit_bin_TRUE-hit_bin_FALSE)

E4.RT_by_hit_by_judgment_wide <- E4.RT_by_hit_by_judgment %>%
  dplyr::select(judge_id,judged,hit_diff) %>%
  spread(judged,hit_diff)%>%
  mutate(diff=pretend-nonpretend)

E4.RT_by_lead1_by_judgment <- E4.click_log_by_judgment %>%
  group_by(judge_id,judged, lead1)%>%
  summarise(RT=median(RT))  %>%
  spread(lead1,RT,sep='_') %>%
  mutate(lead1_diff = lead1_TRUE-lead1_FALSE);

E4.RT_by_lead1_by_judgment_wide <- E4.RT_by_lead1_by_judgment %>%
  dplyr::select(judge_id,judged,lead1_diff) %>%
  spread(judged,lead1_diff)%>%
  mutate(diff=pretend-nonpretend)

E4.RT_by_hit_lags_by_judgment <-E4.RT_by_lag1_by_judgment %>%
  merge(E4.RT_by_hit_by_judgment) %>%
  merge(E4.RT_by_lead1_by_judgment)

E4.RT_by_hit_lags_long_by_judgment <- E4.RT_by_hit_lags_by_judgment %>%
  gather("position","difference",c("lag1_diff", "hit_diff", "lead1_diff"))%>%
  mutate(position = ifelse(position=="lag1_diff",1,
                                         ifelse(position=="hit_diff",0,-1))) %>%
  dplyr::select(judge_id,judged,position,difference);

p<- E4.RT_by_hit_lags_long_by_judgment %>% 
  filter(position<2 & position > -2)%>%
  group_by(judged,position) %>%
  summarise(median_difference=median(difference, na.rm=T),
            sem_difference=bootstrap_error(difference,N_perm)) %>%
  ggplot(aes(x=position,color=judged,fill=judged,y=median_difference)) +
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  scale_color_manual(values=c("#69b3a2", "#404080")) +
  scale_x_continuous(breaks=-1:1,labels=c('hit on next trial', 'hit on current trial', 'hit on previous trial'))+
  geom_abline(slope=0,intercept=0,linetype=3)+
  geom_path(size=1)+
  geom_ribbon(aes(ymin=median_difference-sem_difference,ymax=median_difference+sem_difference),alpha=0.3,show.legend=F)+
  labs(fill=' ',x='', y='RT effect (ms) of hit on trial t')+
  theme_classic()

ggsave('figures/judgment/E4_RT_by_lag.png',p,width=4, height=3.5, dpi=300)

p

```

No difference in the effect of current hit on RT between games that were judged to be pretend vs. non-pretend (`r apa_print(E2.RT_by_hit_by_judgment_wide$diff%>%t.test())$full_result`) in Battleship. Similar in Hangman (`r apa_print(E4.RT_by_hit_by_judgment_wide$diff%>%t.test())$full_result`).

No difference in the effect of previous guess outcome on RT between games that were judged to be pretend vs. non-pretend (`r apa_print(E2.RT_by_lag1_by_judgment_wide$diff%>%t.test())$full_result`) in Battleship. Similar in Hangman (`r apa_print(E4.RT_by_lag1_by_judgment_wide$diff%>%t.test())$full_result`).

No difference in the effect of next guess outcome on RT between games that were judged to be pretend vs. non-pretend (`r apa_print(E2.RT_by_lead1_by_judgment_wide$diff%>%t.test())$full_result`) in Battleship. Similar in Hangman (`r apa_print(E4.RT_by_lead1_by_judgment_wide$diff%>%t.test())$full_result`).

### Model based analysis

```{r battleship-p_click, echo = FALSE, cache=TRUE}

E2.mean_P_click_rank_by_judgment <- E2.click_log_with_boards_by_judgment %>%
  group_by(judge_id,judged) %>%
  summarise(p_click_rank=mean(p_click_rank));

E2.mean_P_click_rank_misses_by_judgment <- E2.click_log_with_boards_by_judgment %>%
  rowwise() %>%
  filter(!hit_bin) %>%
  group_by(judge_id,judged) %>%
  summarise(p_click_rank=mean(p_click_rank));

p<- E2.mean_P_click_rank_by_judgment %>%
    ggplot(aes(x=p_click_rank,fill=judged))+
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
    labs(x='p(ship in clicked square) - rank', y='number of players')+
    theme(legend.position=c(0.10,0.85)) +
    # facet_wrap(~genuine_first, nrow=2)+
    # scale_x_reverse() +
    theme_classic();

ggsave('figures/judgment/E2_pclick_rank.pdf',p)
ggsave('figures/judgement/E2_pclick_rank.png',p,width=5,height=2,dpi=300)

p<- E2.mean_P_click_rank_misses_by_judgment %>%
    ggplot(aes(x=p_click_rank,fill=judged))+
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
    labs(x='p(ship in clicked square) - rank', y='number of players')+
    theme(legend.position=c(0.10,0.85)) +
    # facet_wrap(~genuine_first, nrow=2)+
    # scale_x_reverse() +
    theme_classic();

ggsave('figures/judgment/E2_pclick_rank_misses.pdf',p)
ggsave('figures/judgement/E2_pclick_rank_misses.png',p,width=5,height=2,dpi=300)

E2.num_irrational_clicks_by_judgment <- E2.click_log_with_boards_by_judgment %>%
  group_by(judge_id,judged) %>%
  summarise(nic=sum(p_click==0)/5);

p<- E2.num_irrational_clicks_by_judgment %>%
    ggplot(aes(x=nic,fill=judged))+
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    labs(x='mean number of irrational clicks per game', y='number of players')+
    theme(legend.position=c(0.10,0.85)) +
    theme_classic();

ggsave('figures/judgement/E2_num_irrational_clicks.png',p,width=5,height=3,dpi=300)

```

Seems like people are judging irrational games as pretend much more.

### Accuracy

```{r accuracy, echo = FALSE, cache=TRUE}

E2.judgment_summary <- E2.click_log_by_judgment %>% 
  group_by(subj_id,grid_number) %>%
  summarise(judged=as.factor(judged[1]),test_part=as.factor(test_part[1]))

table(E2.judgment_summary$test_part,E2.judgment_summary$judged)

E4.judgment_summary <- E4.click_log_by_judgment %>% 
  group_by(subj_id,word) %>%
  summarise(judged=as.factor(judged[1]),test_part=as.factor(test_part[1]))

table(E4.judgment_summary$test_part,E4.judgment_summary$judged)

```